Controlling the rheological properties of well-defined polyelectrolytes hydrogels by fine tuning their ion pairing process

### 2015-12-21 PALSE/ion_pairing_paper.tex
fig:strain, as from python script, except switch b and c and no dashed lines

\documentclass[journal=jacsat,manuscript=article]{achemso}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amssymb}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage{siunitx}

@@

Rheological measurements. In order to refine these visual observations, we performed rheological studies of the water-swelling materials. Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}.15, 16 The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process. Excess material was removed, so that the sample exactly filled the gap. Due to Weissenberg effect at high shear rate, no preshear was applied before measurement. To minimize water absorption, the geometry was covered with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap. After one minute of equilibration we performed an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$). For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile was that of pure water which confirms the visual observation. However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G"$) at small strain and flow at large strains ($G" \gg G^\prime$). Solidlike behavior at low strain was confirmed for all accessible frequencies (see supplementary material S6 Fig. 1). We define the critical strain value $\gamma_c$ between these two regimes when $G"$ reaches a maximum. We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest. %To understand the influence of the ion pairing on the mechanical properties, we compute the cohesive energy (equation 1) that should probe the strength of the physical supramolecular bonds.15,16

\begin{figure}
\caption{Storage modulus $G^\prime$ (filled symbols and solid lines) and loss modulus $G"$ (open symbols and dashed lines) measured through oscillatory shear experiments for (a) \ce{PPyr+Cl-} ($\bullet$), \ce{PPyr+Br-} ($\blacksquare$) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+Br-} ($\blacklozenge$) and \ce{PPyr+Br-} ($\blacksquare$) and (c) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$). $G^\prime$  and $G"$ measured during an amplitude sweep at a fixed frequency $f=\SI{1}{\hertz}$ are plotted against the strain amplitude $\gamma$. All samples are at 8\%wt }
\label{fig:strain}
\end{figure}

@@

We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c. For both bromide (Figure~\ref{fig:strain}b, X=Br) and iodide (Figure~\ref{fig:strain}c, X=I), imidazolium based gels are much weaker than their pyrrolidinium counterpart.

>>>>

\documentclass[journal=jacsat,manuscript=article]{achemso}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amssymb}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage{siunitx}
\usepackage{xcolor}

@@

Rheological measurements. In order to refine these visual observations, we performed rheological studies of the water-swelling materials. Briefly, data were recorded at \SI{25}{\degree} with an AR 1000 rheometer (TA Instruments) in a cone-plate geometry of radius \SI{40}{\milli\metre}, an angle of \SI{2}{\degree} and a truncation of \SI{58}{\micro\metre}.15, 16 The sample was placed on the plate, the cone was then lowered to the measuring position, spreading the sample in the process. Excess material was removed, so that the sample exactly filled the gap. Due to Weissenberg effect at high shear rate, no preshear was applied before measurement. To minimize water absorption, the geometry was covered with a solvent trap, using light mineral oil as a liquid seal between the rotor and the cap. After one minute of equilibration we performed an oscillatory frequency sweep at small amplitude (strain amplitude $\gamma=0.1\%$) and then an oscillatory strain sweep at fixed frequency ($f=\SI{1}{\hertz}$). For fluoride counterions as well as \ce{PIm+Cl-}, the rheological profile was that of pure water which confirms the visual observation. However as shown in Figure~\ref{fig:strain}, all other samples are solidlike ($G^\prime \gg G^{\prime\prime}$) at small strain and flow at large strains ($G^{\prime\prime} \gg G^\prime$). Solidlike behavior at low strain was confirmed for all accessible frequencies (see supplementary material S6 Fig. 1). We define the critical strain value $\gamma_c$ between these two regimes when $G^{\prime\prime}$ reaches a maximum. We checked that the linear mechanical properties are unchanged by the flow history given a few minutes of rest. %To understand the influence of the ion pairing on the mechanical properties, we compute the cohesive energy (equation 1) that should probe the strength of the physical supramolecular bonds.15,16

\begin{figure}
\includegraphics{strain.pdf}
\caption{Storage modulus $G^\prime$ (filled symbols) and loss modulus $G^{\prime\prime}$ (open symbols) measured through oscillatory shear experiments for (a) \ce{PPyr+Cl-} (\textcolor{lightgray}{$\bullet$}), \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}) and \ce{PPyr+I-} ($\blacktriangle$), (b) \ce{PIm+I-} ($\blacktriangledown$) and \ce{PPyr+I-} ($\blacktriangle$) and (c) \ce{PIm+Br-} (\textcolor{gray}{$\blacklozenge$}) and \ce{PPyr+Br-} (\textcolor{gray}{$\blacksquare$}). $G^\prime$  and $G^{\prime\prime}$ measured during an amplitude sweep at a fixed frequency $f=\SI{1}{\hertz}$ are plotted against the strain amplitude $\gamma$. All samples are at 8\%wt }
\label{fig:strain}
\end{figure}

@@

We also compare the rheology of \ce{PIm+X-} and \ce{PPyr+X-} with the same counterion \ce{X} in Figure~\ref{fig:strain}b,c. For both iodide (Figure~\ref{fig:strain}b, X=I) and bromide (Figure~\ref{fig:strain}c, X=Br), imidazolium based gels are much weaker than their pyrrolidinium counterpart.
